### 2023-12-04 Appendix.Rmd
Add appendix

# 1. Interaction effects of inter-factor correlation and factor loadings 
```{r}
appendix_results <- here::here("appendix_results.zip")
if(!fs::dir_exists(fs::path_ext_remove(appendix_results))){
  if(!fs::file_exists(appendix_results))
    download.file("https://keeper.mpdl.mpg.de/f/7cee7f676c0d4a0a8b80/?dl=1", appendix_results, mode = "wb")
  unzip(appendix_results, exdir=here::here("appendix_results"))
}
```

```{r echo = FALSE, message = FALSE, fig.width=11.3, fig.height=8.6, fig.align = 'center',fig.cap = "Estimated statistical power for recovering the data-generating model (y-axis) as a function of sample size (x-axis) depending on (A) loading strength for lower versus higher inter-factor correlation and (B) inter-factor correlation for lower versus higher factor loadings. * denotes results from the original analyses in the main manuscript"}

theme_plots <- theme(
  panel.background = element_blank(),
  panel.grid.major = element_line(color = "grey95"),
  legend.key       = element_rect(fill = NA),
  axis.line        = element_line(
    colour = "black",
    linewidth = 0.5,
    linetype = "solid"
  ),
  axis.text.x      = element_text(
    size = 16,
    family = "sans",
    hjust = 0.5,
    vjust = -0.5
  ),
  axis.text.y      = element_text(
    size = 16,
    family = "sans",
    hjust = 0,
    vjust = 0.5
  ),
  axis.title.x     = element_text(
    size = 20,
    hjust = 0.5,
    vjust = -0.5
  ),
  axis.title.y     = element_text(
    size = 20,
    hjust = -0.1,
    vjust = 0.5
  ),
  plot.title       = element_text(
    lineheight = .8,
    face = "bold",
    size = 22
  ),
  legend.title     = element_text(size = 20),
  legend.text      = element_text(size = 18),
  legend.position  = "right"
)

results_load_model3 <- 
  read.csv(here::here("appendix_results", "results_load_model3.csv"))
results_cov_model3 <- 
  read.csv(here::here("appendix_results", "results_cov_model3.csv"))
results_load_model3_highcov <- 
  read.csv(here::here("appendix_results", "results_load_model3_highcov.csv"))
results_cov_model3_lowload <- 
  read.csv(here::here("appendix_results", "results_cov_model3_lowload.csv"))

plot_load_model3 <- results_load_model3 %>%
  ggplot(aes(sample_size, best_3 * 100, colour = factor(loading_strength))) +
  geom_point() +
  geom_line(alpha = 0.4,
            linewidth = 1.3,
            key_glyph = "rect") +
  geom_hline(
    yintercept = 95,
    linetype = "dashed",
    color = "#F29090",
    linewidth = 1.3
  ) +
  xlab('Sample size\n') + ylab('Power (%)') + 
  ggtitle('Inter-factor correlation = 0.3* \n') +
  scale_color_discrete('Loading       \nstrength') +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(50, 200),
    breaks = seq(50, 200, by = 25)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 100),
    breaks = seq(0, 100, by = 20)
  ) +
  theme_plots

plot_load_model3_highcov <- results_load_model3_highcov %>%
  ggplot(aes(sample_size, best_3 * 100, colour = factor(loading_strength))) +
  geom_point() +
  geom_line(alpha = 0.4,
            linewidth = 1.3,
            key_glyph = "rect") +
  geom_hline(
    yintercept = 95,
    linetype = "dashed",
    color = "#F29090",
    linewidth = 1.3
  ) +
  xlab('Sample size\n') + ylab('Power (%)') + ggtitle('Inter-factor correlation = 0.7\n') +
  scale_color_discrete('Loading \nstrength') +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(50, 200),
    breaks = seq(50, 200, by = 25)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 100),
    breaks = seq(0, 100, by = 20)
  ) +
  theme_plots

plot_cov_model3_lowload <- results_cov_model3_lowload %>%
  ggplot(aes(sample_size, best_3 * 100, colour = factor(cov1))) +
  geom_point() +
  geom_line(alpha = 0.4,
            linewidth = 1.3,
            key_glyph = "rect") +
  geom_hline(
    yintercept = 95,
    linetype = "dashed",
    color = "#F29090",
    linewidth = 1.3
  ) +
  xlab('Sample size\n') + ylab('Power (%)') + 
  ggtitle('Factor loading = 0.5\n') +
  scale_color_discrete('Inter-factor \ncorrelation') +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(50, 200),
    breaks = seq(50, 200, by = 25)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 100),
    breaks = seq(0, 100, by = 20)
  ) +
  theme_plots

plot_cov_model3 <- results_cov_model3 %>%
  ggplot(aes(sample_size, best_3 * 100, colour = factor(cov1))) +
  geom_point() +
  geom_line(alpha = 0.4,
            linewidth = 1.3,
            key_glyph = "rect") +
  geom_hline(
    yintercept = 95,
    linetype = "dashed",
    color = "#F29090",
    linewidth = 1.3
  ) +
  xlab('Sample size\n') + ylab('Power (%)') + 
  ggtitle('Factor loading = 0.7*\n') +
  scale_color_discrete('Inter-factor \ncorrelation') +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(50, 200),
    breaks = seq(50, 200, by = 25)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 100),
    breaks = seq(0, 100, by = 20)
  ) +
  theme_plots

joined_plot <-
  ggarrange(plot_cov_model3_lowload,
            plot_cov_model3,
            plot_load_model3,
            plot_load_model3_highcov,
            ncol = 2,
            nrow = 2)
joined_plot

```